Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-02-13 ms.tex
Added some margin notes

If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
%
\begin{eqnarray*}
    \mathcal{L}(
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
        \mathbf{n}, t_\oplus, D_\mathrm{L}, x, \psi) \\
    &= \mathcal{L}_\mathrm{TOA}(
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
        &\times
        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
         \mathbf{n}, D_\mathrm{L}, x, \psi
    ).
\end{eqnarray*}
%
The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{TOA} \propto
    \exp \left[
            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
        \right]
    = \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
    \right]
\end{equation}
%
The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $x = \cos \iota$, and polarization angle $\psi$.
%
\begin{equation}
    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, x, \psi) \right)^2
    \right]
\end{equation}
%
It is easily shown that
%
\begin{eqnarray*}
    {\rho_i}^2 &=
        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left[
            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
            \left(1 + 6 x^2 + x^4 \right) \\
            &+ \left(1 - x^2 \right) \left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi}
        \right],
\end{eqnarray*}
%
where $F_{+,i}$ and $F_{\times,i}$ are the `$+$' and `$\times$' antenna factors of detector $i$, themselves functions of sky location $\mathbf{n}$.

@@

The innermost integration is over $D_\mathrm{L}$ and is performed using adaptive Gaussian quadrature, but with a twist that speeds up convergence.

@@

>>>>

\marginpar{Explain why none of these expressions depend on the coalescence phase $\phi_c$.}%
%
If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
%
\begin{eqnarray*}
    \mathcal{L}(
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
        \mathbf{n}, t_\oplus, D_\mathrm{L}, x, \psi) \\
    &= \mathcal{L}_\mathrm{TOA}(
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
        &\times
        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
         \mathbf{n}, D_\mathrm{L}, x, \psi
    ).
\end{eqnarray*}
%
The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{TOA} \propto
    \exp \left[
            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
        \right]
    = \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
    \right]
\end{equation}
%
The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $x = \cos \iota$, and polarization angle $\psi$.
%
\begin{equation}
    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, x, \psi) \right)^2
    \right]
\end{equation}
%
\marginpar{Cite \ac{GRB} prospects paper appendix; a very similar equation appears there. Is there somewhere else that has this formula? Shutz doesn't, and Finn \& Chernoff's notation is too heavy.}%
%
It is easily shown that
%
\begin{eqnarray*}
    {\rho_i}^2 &=
        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left[
            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
            \left(1 + 6 x^2 + x^4 \right) \\
            &+ \left(1 - x^2 \right) \left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi}
        \right],
\end{eqnarray*}
%
where $F_{+,i}$ and $F_{\times,i}$ are the `$+$' and `$\times$' antenna factors of detector $i$, themselves functions of sky location $\mathbf{n}$.

@@

\marginpar{Explain initial integration breakpoints and how they are calculated. Cite \texttt{gsl\_integrate\_qagp}.}%
%
The innermost integration is over $D_\mathrm{L}$ and is performed using adaptive Gaussian quadrature, but with a twist that speeds up convergence.

@@

\section{Hierarchical refinement}

\marginpar{Talk about adaptive selection of HEALPix resolution and of pixels on which to evaluate \ac{SNR} marginal posterior. Something to think about: this is logically easier to explain last, but chronologically makes more sense first.}
